---
title: "Assessment 10 - Data Visualization Principles Part 2"
author: "Alessandro Corradini - Harvard Data Science Professional"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br/>

## **Customizing plots - watch and learn**
To make the plot on the right in the exercise from the last set of assessments, we had to reorder the levels of the states' variables.

**Instructions**

- Redefine the ```state``` object so that the levels are re-ordered by rate.
- Print the new object ```state``` and its levels so you can see that the vector is now re-ordered by the levels.

```{r, include=TRUE}
state <- reorder(state, rate)
print(state)
levels(state)
```

<br/>

## **Customizing plots - redefining**
Now we are going to customize this plot a little more by creating a rate variable and reordering by that variable instead.

**Instructions**

- Add a single line of code to the definition of the ```dat``` table that uses ```mutate``` to reorder the states by the rate variable.
- The sample code provided will then create a bar plot using the newly defined ```dat```.

```{r,include=TRUE}
dat <- us_contagious_diseases %>% filter(year == 1967 & disease=="Measles" & count>0 & !is.na(population)) %>%
 mutate(rate = count / population * 10000 * 52 / weeks_reporting) %>%
 mutate(state = reorder(state, rate))
dat %>% ggplot(aes(state, rate)) +
 geom_bar(stat="identity") +
 coord_flip()
```

<br/>

## **Showing the data and customizing plots**
Say we are interested in comparing gun homicide rates across regions of the US. We see this plot:

```{r, include=TRUE}
library(dplyr)
library(ggplot2)
library(dslabs)
data("murders")
murders %>% mutate(rate = total/population*100000) %>%
  group_by(region) %>%
  summarize(avg = mean(rate)) %>%
  mutate(region = factor(region)) %>%
  ggplot(aes(region, avg)) +
  geom_bar(stat="identity") +
  ylab("Murder Rate Average")
```

and decide to move to a state in the western region. What is the main problem with this interpretaion?


**Instructions**

Possible Answers

- The categories are ordered alphabetically.
- The graph does not show standard errors.
I- t does not show all the data. We do not see the variability within a region and it's possible that the safest states are not in the West. [X]
- The Northeast has the lowest average.


## **Making a box plot**
To further investigate whether moving to the western region is a wise decision, let's make a box plot of murder rates by region, showing all points.

**Instructions**

- Make a box plot of the murder rates by region.
- Order the regions by their median murder rate.
- Show all of the points on the box plot.

```{r, include=TRUE}
library(dplyr)
library(ggplot2)
library(dslabs)
data("murders")
murders %>% mutate(rate = total/population*100000) %>%
  mutate(region=reorder(region, rate, FUN=median)) %>%
  ggplot(aes(region, rate)) +
  geom_boxplot() +
  geom_point()
```  