---
title: "Assessment 01 - R Basics"
author: "Alessandro Corradini - Harvard Data Science Professional"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br/>

### **Using variables 1**
What is the sum of the first n positive integers?
We can use the formula $\ n(n+1)/2$ to quickly compute this quantity.

**Instructions**

- Define 'n=100'
- Then use R to compute the sum of 1 through 100 using the formula $\ n(n+1)/2$. What is the sum?

```{r, include=TRUE}
# Here is how you compute the sum for the first 20 integers
20*(20+1)/2 

# However, we can define a variable to use the formula for other values of n
n <- 20
n*(n+1)/2

n <- 25
n*(n+1)/2

# Below, write code to calculate the sum of the first 100 integers
n <- 100
n*(n+1)/2
```

<br/>

### **Using variables 2**
What is the sum of the first 1000 positive integers?

We can use the formula $\ n(n+1)/2$ to quickly compute this quantity.

**Instructions**

- Use the same formula as the last exercise but substitute n
- Instead of typing the result, use the formula and defined variable.

```{r, include=TRUE}
# Below, write code to calculate the sum of the first 1000 integers 
n <- 1000
n*(n+1)/2
```

<br/>

### **Functions**
Run the following code in the R console:

```
n <- 1000
x <- seq(1,n)
sum(x)
```
Based on the result, what do you think the functions ```seq``` and ```sum``` do? You can use the help system.

**Possible Answers**

- sum creates a list of numbers and seq adds them up.
- seq creates a list of numbers and sum adds them up. [X]
- seq computes the difference between two arguments and sum computes the sum of 1 through 1000.
- sum always returns the same number

<br/>

### **Nested function calls 1**
In math and programming we say we evaluate a function when we replace arguments with specific values. So if we type ```log2(16)``` we evaluate the ```log2``` function to get the log to the base 2 of 16 which is 4.

In R it is often useful to evaluate a function inside another function. For example, ```sqrt(log2(16))``` will calculate the log to the base 2 of 16 and then compute the square root of that value. So the first evaluation gives a 4 and this gets evaluated by sqrt to give the final answer of 2.

**Instructions**

- Use one line of code to compute the log, to the base 10, of the square root of 100.
- Make sure your code includes the ```log10``` and ```sqrt``` functions.

```{r, include=TRUE}
# log to the base 2 
log2(16)

# sqrt of the log to the base 2 of 16:
sqrt(log2(16))

# Compute log to the base 10 (log10) of the sqrt of 100. Do not use variables.
log10(sqrt(100))
```

<br/>

### **Nested functions call 2**
Which of the following will always return the numeric value stored in x? You can try out examples and use the help system in the R console.

**Possible Answers**

- ```log(10^x)```
- ```log10(x^10)```
- ```log(exp(x))``` [X]
- ```exp(log(x, base = 2))```