---
title: "Assessment 08 - Data Types"
author: "Alessandro Corradini - Harvard Data Science Professional"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br/>

## **Comparing Proportions of Hits**
In a previous exercise, we determined whether or not each poll predicted the correct winner for their state in the 2016 U.S. presidential election. Each poll was also assigned a grade by the poll aggregator. Now we're going to determine if polls rated A- made better predictions than polls rated C-.

In this exercise, filter the errors data for just polls with grades A- and C-. Calculate the proportion of times each grade of poll predicted the correct winner.

Instructions

- Filter errors for grades A- and C-.
- Group the data by grade and hit.
- Summarize the number of hits for each grade.
- Generate a two-by-two table containing the number of hits and misses for each grade.
- Calculate the proportion of times each grade was correct.

```{r, include=TRUE}
# The 'errors' data have already been loaded. Examine them using the `head` function.
head(errors)

# Generate an object called 'totals' that contains the numbers of good and bad predictions for polls rated A- and C-
totals <- errors %>%
  filter(grade %in% c("A-", "C-")) %>%
  group_by(grade,hit) %>%
  summarize(num = n()) %>%
  spread(grade, num)

# Print the proportion of hits for grade A- polls to the console
totals[[2,3]]/sum(totals[[3]])

# Print the proportion of hits for grade C- polls to the console
totals[[2,2]]/sum(totals[[2]])
```

<br/>

## **Chi-squared Test**
We found that the A- polls predicted the correct winner about 86% of the time in their states and C- polls predicted the correct winner about 80% of the time.

Use a chi-squared test to determine if these proportions are different.

Instructions

- Use the ```chisq.test``` function to perform the chi-squared test. Save the results to an object called ```chisq_test```.
- Print the p-value of the test to the console.


```{r}
# The 'totals' data have already been loaded. Examine them using the `head` function.
head(totals)

# Perform a chi-squared test on the hit data. Save the results as an object called 'chisq_test'.
chisq_test <- totals %>% 
  select(-hit) %>%
  chisq.test()
chisq_test

# Print the p-value of the chi-squared test to the console
chisq_test$p.value
```

<br/>

## **Odds Ratio Calculation**
It doesn't look like the grade A- polls performed significantly differently than the grade C- polls in their states.

Calculate the odds ratio to determine the magnitude of the difference in performance between these two grades of polls.

Instructions

- Calculate the odds that a grade C- poll predicts the correct winner. Save this result to a variable called ```odds_C```.
- Calculate the odds that a grade A- poll predicts the correct winner. Save this result to a variable called ```odds_A```.
-Calculate the odds ratio that tells us how many times larger the odds of a grade A- poll is at predicting the winner than a grade C- poll.

```{r, include=TRUE}
# The 'totals' data have already been loaded. Examine them using the `head` function.
head(totals)

# Generate a variable called `odds_C` that contains the odds of getting the prediction right for grade C- polls
odds_C <- (totals[[2,2]] / sum(totals[[2]])) / 
  (totals[[1,2]] / sum(totals[[2]]))

# Generate a variable called `odds_A` that contains the odds of getting the prediction right for grade A- polls
odds_A <- (totals[[2,3]] / sum(totals[[3]])) / 
  (totals[[1,3]] / sum(totals[[3]]))

# Calculate the odds ratio to determine how many times larger the odds ratio is for grade A- polls than grade C- polls
odds_A/odds_C
```

<br/>

## **Significance**
We did not find meaningful differences between the poll results from grade A- and grade C- polls in this subset of the data, which only contains polls for about a week before the election. Imagine we expanded our analysis to include all election polls and we repeat our analysis. In this hypothetical scenario, we get that the p-value for the difference in prediction success if 0.0015 and the odds ratio describing the effect size of the performance of grade A- over grade B- polls is 1.07.

Based on what we learned in the last section, which statement reflects the best interpretation of this result?

Possible Answers

- The p-value is below 0.05, so there is a significant difference. Grade A- polls are significantly better at predicting winners.
- The p-value is too close to 0.05 to call this a significant difference. We do not observe a difference in performance.
- The p-value is below 0.05, but the odds ratio is very close to 1. There is not a scientifically significant difference in performance. [X]
- The p-value is below 0.05 and the odds ratio indicates that grade A- polls perform significantly better than grade C- polls.