---
title: "Assessment 07 - The t-distribution"
author: "Alessandro Corradini - Harvard Data Science Professional"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br/>

## **Using the t-Distribution**
We know that, with a normal distribution, only 5% of values are more than 2 standard deviations away from the mean.

Calculate the probability of seeing t-distributed random variables being more than 2 in absolute value when the degrees of freedom are 3.

Instructions

Use the ```pt``` function to calculate the probability of seeing a value less than or equal to the argument.

```{r, include=TRUE}
# Calculate the probability of seeing t-distributed random variables being more than 2 in absolute value when 'df = 3'.
1 - pt(2, 3) + pt(-2, 3)
```

<br/>

## **Plotting the t-distribution**
Now use sapply to compute the same probability for degrees of freedom from 3 to 50.

Make a plot and notice when this probability converges to the normal distribution's 5%.

Instructions

- Make a vector called ```df``` that contains a sequence of numbers from 3 to 50.
- Using ```function```, make a function called ```pt_func``` that recreates the calculation for the probability that a value is greater than 2 as an absolute value for any given degrees of freedom.
- Use ```sapply``` to apply the ```pt_func``` function across all values contained in ```df```. Call these probabilities ```probs```.
- Use the ```plot``` function to plot ```df``` on the x-axis and ```probs``` on the y-axis.


```{r}
# Generate a vector 'df' that contains a sequence of numbers from 3 to 50
df <- seq(3,50)

# Make a function called 'pt_func' that calculates the probability that a value is more than |2| for any degrees of freedom 
pt_func <- function(n) {
  1 - pt(2, n) + pt(-2, n)
}

# Generate a vector 'probs' that uses the `pt_func` function to calculate the probabilities
probs <- sapply(df, pt_func)

# Plot 'df' on the x-axis and 'probs' on the y-axis
plot(df, probs)
```

<br/>

## **Sampling From the Normal Distribution**
In a previous section, we repeatedly took random samples of 50 heights from a distribution of heights. We noticed that about 95% of the samples had confidence intervals spanning the true population mean.

Re-do this Monte Carlo simulation, but now instead of N=50, use N=15. Notice what happens to the proportion of hits.

Instructions

- Use the ```replicate``` function to carry out the simulation. Specify the number of times you want the code to run and, within brackets, the three lines of code that should run.
- First use the ```sample``` function to randomly sample ```N``` values from ```x```.
- Second, create a vector called i```nterval``` that calculates the 95% confidence interval for the sample. You will use the ```qnorm``` function.
- Third, use the ```between``` function to determine if the population mean ```mu``` is contained between the confidence intervals.
- Save the results of the Monte Carlo function to a vector called ```res```.
- Use the ```mean``` function to determine the proportion of hits in ```res```.

```{r, include=TRUE}
# Load the neccessary libraries and data
library(dslabs)
library(dplyr)
data(heights)

# Use the sample code to generate 'x', a vector of male heights
x <- heights %>% filter(sex == "Male") %>%
  .$height

# Create variables for the mean height 'mu', the sample size 'N', and the number of times the simulation should run 'B'
mu <- mean(x)
N <- 15
B <- 10000

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling
set.seed(1)

# Generate a logical vector 'res' that contains the results of the simulations
res <- replicate(B, {
  s <- sample(x, N, replace = TRUE)
  interval <- c(qnorm(0.025, mean(s), sd(s)/sqrt(N)), qnorm(0.975, mean(s), sd(s)/sqrt(N)))
  between(mu, interval[1], interval[2])
})

# Calculate the proportion of times the simulation produced values within the 95% confidence interval. Print this value to the console.
mean(res)
```

<br/>

## **Sampling from the t-Distribution**
N=15 is not that big. We know that heights are normally distributed, so the t-distribution should apply. Repeat the previous Monte Carlo simulation using the t-distribution instead of using the normal distribution to construct the confidence intervals.

What are the proportion of 95% confidence intervals that span the actual mean height now?

Instructions

- Use the ```replicate``` function to carry out the simulation. Specify the number of times you want the code to run and, within brackets, the three lines of code that should run.
- First use the ```sample``` function to randomly sample ```N``` values from ```x```.
- Second, create a vector called ```interval``` that calculates the 95% confidence interval for the sample. Remember to use the ```qt``` function this time to generate the confidence interval.
- Third, use the ```between``` function to determine if the population mean ```mu``` is contained between the confidence intervals.
- Save the results of the Monte Carlo function to a vector called ```res```.
- Use the ```mean``` function to determine the proportion of hits in ```res```.

```{r, include=TRUE}
# The vector of filtered heights 'x' has already been loaded for you. Calculate the mean.
mu <- mean(x)

# Use the same sampling parameters as in the previous exercise.
set.seed(1)
N <- 15
B <- 10000

# Generate a logical vector 'res' that contains the results of the simulations using the t-distribution
res <- replicate(B, {
  s <- sample(x, N, replace = TRUE)
  interval <- c(mean(s) - qt(0.975, N - 1) * sd(s) / sqrt(N), mean(s) + qt(0.975, N - 1) * sd(s) / sqrt(N))
  between(mu, interval[1], interval[2])
})

# Calculate the proportion of times the simulation produced values within the 95% confidence interval. Print this value to the console.
mean(res)
```

<br/>

## **Why the t-Distribution?**
Why did the t-distribution confidence intervals work so much better?

Possible Answers

- The t-distribution takes the variability into account and generates larger confidence intervals. [X]
- Because the t-distribution shifts the intervals in the direction towards the actual mean.
- This was just a chance occurrence. If we run it again, the CLT will work better.
- The t-distribution is always a better approximation than the normal distribution.
